Interisland movements and oceanic swimming of woodland caribou in Newfoundland

### 2020-03-18 paper/manuscript.Rmd
n ids/events

The Fogo Island archipelago, off the coast of Newfoundland, Canada, is home to
approximately 300 woodland caribou. During routine fieldwork on May 30, 2017, we
observed an unmarked adult male caribou swim between Western and Eastern Indian
Islands (Figure 1), a distance of at least 470 m which took ~9 minutes (~52 m
per minute). This observation prompted us to investigate whether swimming
behaviour was common among GPS radio-collared caribou in the population (n = 29
adult females; 1240g, GPS 4400M; Lotek Wireless Inc., Newmarket, Ontario,
Canada). We identified swimming events as two consecutive GPS (2-hour relocation
rates) locations from an individual occurred on different islands. Northeastern
Newfoundland typically experiences pack ice during winter and caribou are known
locally to travel between islands by walking over the ice. We restricted our GPS
data to the ice-free period of the year (April 1 - December 31) and we
identified  adult female caribou that collectively had `r nids` swimming events
over three years (Figure 2). In addition to our own visual and remotely sensed
observations, residents of Fogo Island have also reported observing caribou
swimming between islands on numerous occasions.

@@

In total, 12 of 29 collared female caribou swam between islands in the Fogo
archipelago. Only 3 of these 12 had more than XX swimming events during the
study period; ~30% of the marked population swam at least XX times, whereas for
~10% of individuals these swimming events were relatively common, on average
occurring once every XX  days during ice-free seasons. We did not
detect any swimming events for the remaining 17 individuals. It remains unclear
whether these individuals do not swim, or whether swimming is rare for most caribou. 
We are also unable to draw conclusions about the prevalence of
swimming in the ocean by male caribou: our initial visual observation was of an
adult male, but our GPS data only included females.

>>>>

The Fogo Island archipelago, off the coast of Newfoundland, Canada, is home to
approximately 300 woodland caribou. During routine fieldwork on May 30, 2017, we
observed an unmarked adult male caribou swim between Western and Eastern Indian
Islands (Figure 1), a distance of at least 470 m which took ~9 minutes (~52 m
per minute). This observation prompted us to investigate whether swimming
behaviour was common among GPS radio-collared caribou in the population (n = 29
adult females; 1240g, GPS 4400M; Lotek Wireless Inc., Newmarket, Ontario,
Canada). We identified swimming events as two consecutive GPS (2-hour relocation
rates) locations from an individual occurred on different islands. Northeastern
Newfoundland typically experiences pack ice during winter and caribou are known
locally to travel between islands by walking over the ice. We restricted our GPS
data to the ice-free period of the year (April 1 - December 31) and we
identified `r nids` adult female caribou that collectively had `r nevents` swimming events
over three years (Figure 2). In addition to our own visual and remotely sensed
observations, residents of Fogo Island have also reported observing caribou
swimming between islands on numerous occasions.

@@

In total, 12 of 29 collared female caribou swam between islands in the Fogo
archipelago. Only 3 of these 12 had more than XX swimming events during the
study period; ~30% of the marked population swam at least once, whereas for
~10% of individuals these swimming events were relatively common, on average
occurring once every XX  days during ice-free seasons. We did not
detect any swimming events for the remaining 17 individuals. It remains unclear
whether these individuals do not swim, or whether swimming is rare for most caribou. 
We are also unable to draw conclusions about the prevalence of
swimming in the ocean by male caribou: our initial visual observation was of an
adult male, but our GPS data only included females.
